---
title       : Portfolio Risk Visualizer
subtitle    : Illustrating how portfolio weights affect portfolio risk
author      : Thomas M. Anichini
job         : Signature Track student
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
url: 
#  lib: ../../libraries
  lib: ../../librariesNew
  assets: ../../portfolioRiskVis/portfolioRiskVis_slidify/assets
widgets     : [mathjax, quiz, bootstrap]
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---

## Portfolio Risk Visualizer

- "Risk-based" investment strategies have proliferated over the past decade

- Many investors employ these strategies without understanding how they work

- The application [portfolioRiskViz](https://mcarlo.shinyapps.io/portfolioRiskViz/)
advances their understanding by illustrating how a portfolio's risk changes when
its asset weights change

---

## Side panel: play with portfolio allocations

- Sliders invite the user to change allocations among several investments 

+ Assets include

  - Commodities
  - REITs
  - Bonds
  - International Stocks
  - US Stocks

---

## Portfolio Composition panel: Reactive charts

- Resulting changes in risk contributions are nonlinear and therefore not
intuitive

```{r portcalcs,echo=FALSE,results='hide',cache=TRUE}
Asset <- c("US Stocks","Int'l Stocks","Bonds","REITs","Commodities")
weights <- rep(.2, 5)

Volatility <- c(0.2, .25, .05, .25, .25)
Expected_Return <- c(0.085, .08, .02, .075, .05)

volDF <- data.frame(cbind(Volatility, Expected_Return))
rownames(volDF) <- Asset

correlations <- matrix(cbind(c(1, 0.95, 0.15, 0.85, 0.6),
                             c(0.95, 1, 0.3, 0.8, 0.7),
                             c(0.15, 0.3, 1, 0.25, 0.05),
                             c(0.85, 0.8, 0.25, 1, 0.45),
                             c(0.6, 0.7, 0.05, 0.45, 1)), nrow = 5)



colnames(correlations) <- Asset
rownames(correlations) <- Asset
```

```{r riskDemo,echo=TRUE,results='hide'}
covariances <- correlations * (Volatility %*% t(Volatility))

contributions <- weights * (covariances %*% weights)
variance <- sum(contributions)
risk_contributions <- contributions/variance

```

```{r portcomp,echo=FALSE,fig.align='center',fig.height=5}
return_contributions <- weights * Expected_Return

portfolio_std <- sqrt(variance)
portfolio_return <- sum(return_contributions)

axisLabels <- seq(-2., 2., .2)
returnLabels <-  seq(-.10, .10, .01)
rrLabelsX <-  seq(0, 0.35, .05)
rrLabelsY <-  seq(0, 0.10, .01)
    par(mfcol = c(1,3), oma = c(0.5, 5, 0.5, 0.5))
    barplot(t(weights), horiz = TRUE, main = "Weight", ps = 12, cex.main = 1.5, xaxt = "n", col = "#4F81BD", yaxt = "n", mar = c(0.5, 20, 0.5, 0.5) + .1)
    axis(1, at = axisLabels, lab=paste0(axisLabels * 100, " %"), 
         las=TRUE, cex.axis=1, padj = 0.5)
    axis(2, at=c(0.85,2,3.15,4.3,5.45), lab = Asset , las=TRUE, lty = 0, 
         ps = 12, cex.axis = 1.5)
    barplot(t(risk_contributions), 
            main = "Risk Contribution", ps = 12, cex.main = 1.5, col = "#9BBB59", xaxt = "n", 
            yaxt = "n", horiz = TRUE, mar = c(4, 0.5, 0.5, 0.5) + .1)
    axis(1, at = axisLabels, lab=paste0(axisLabels * 100, " %"), 
         las=TRUE, cex.axis=1, padj = 0.5)
    barplot(t(return_contributions), 
            main = "Return Contribution", ps = 12, cex.main = 1.5, col = "#FF9800", xaxt = "n", 
            yaxt = "n", horiz = TRUE, mar = c(4, 0.5, 0.5, 0.5) + .1)
    axis(1, at = returnLabels, lab=paste0(returnLabels * 100, " %"), 
         las=TRUE, cex.axis=1, padj = 0.5)

```

---

## Non-reactive panels illustrate common portfolio strategies

- Non-reactive panels depict Portfolio Composition graphs for Minimum Variance, 
Maximum Sharpe Ratio, and Risk Parity portfolios

- Comments draw user's attention to features of these strategies not commonly
known outside the world of financial academics and quantitative portfolio
managers

- Assumptions panel lists assumptions for asset Volatilities, Expected Returns,
and a correlation matrix

- With these assumptions, a competent investment professional could easily
reproduce these results, whether in R, Excel, or any other platform

- Definitions panel clarifies mathematical definitions of terms used in the app
